---
title: "EARLI EHR Vitals and Labs Clinical Classifier Model"
author: "Manoj Maddali, MD"
date: "7/14/2021"
output: html_document
---

# Setup the model
Set up and train the model using only vitals and labs features.  
**Training set:** ARMA, ALVEOLI, FACTT  
**Vitals and Labs Features:** temperature, systolic BP, HR, RR, vasopressor use, glucose, hematocrit, WBC count, platelet count, sodium, creatinine, albumin, bilirubin, and bicarbonate  

```{r Setup, include=TRUE, echo=TRUE, messages=FALSE}
## Load the required packages
library(caret)
library(dplyr)
library(ggplot2)
library(survival)
library(survminer)
library(ggfortify)
library(ggalluvial)
library(glue)
library(readr)
library(pROC)
library(xgboost)
library(OpenMPController)
library(RColorBrewer)
library(glue)
require(reshape2)

## Set the seed
set.seed(1234)
omp_set_num_threads(8) # caret parallel processing threads

# No scientific notation
options(scipen=999999)

## Load the data
# Training set
training <- read_csv('*****')
training <- as_tibble(training)

# Validation set - EARLI EHR derived
earli_ehr <- read_csv('*****')
earli_ehr <- as_tibble(earli_ehr)

# Validation set - EARLI original, hand-curated
earli_curated <- read_csv('*****')
earli_curated <- as_tibble(earli_curated)
# Select only the patients that are in EARLI EHR
earli_curated <- earli_curated[earli_curated$ID %in% earli_ehr$ID, ]

# Select model features (only vitals and labs)
vitals_labs = c("tempch", "sysbpl", "hrateh", "resph", "vasop",
                "gluch", # not present in VALID
                "hctl", "wbch", "plate", "sodiumh", "creath", "albuml", "bili", "bicar",
                "Class")

# Use vitals and labs only
training = subset(training, select = vitals_labs)

## Set up the hyperparameter search grid for EARLI based on AJRCCM original model (exact same features)
# nrounds = 100, max_depth = 4, eta = 0.1, gamma = 0, colsample_bytree = 1, min_child_weight = 1, subsample = 1
grid_earli <- expand.grid(
 nrounds = 100,
 max_depth = 4,
 eta = 0.1,
 gamma = 0,
 colsample_bytree = 1,
 min_child_weight = 1,
 subsample = 1
)

# Set up 10 fold cross validation training 
train_control <- caret::trainControl(
  method = 'cv',
  number = 10, 
  verboseIter = TRUE,
  allowParallel = TRUE
)

# Train the model using vitals and labs only
xgb <- caret::train(
  Class ~ .,
  data = training,
  trControl = train_control,
  tuneGrid = grid_earli,
  na.action = na.pass,
  method = 'xgbTree',
  verbose = TRUE
)

```
# Evaluate in training set (ARMA, ALVEOLI, FACTT)
```{r Evaluate in training set, include=TRUE, echo=TRUE, messages=TRUE}
# Evaluate the model on the training set
xgb.train = predict(xgb, training, na.action = na.pass, type="prob")
xgb.train = xgb.train[,2] # Get whether or not we are predicting class 1 (hyper)
xgb.train.class = ifelse(xgb.train >= 0.50, 1, 0)
# ROC
xgb.train.roc = roc(predictor = xgb.train, training$Class, levels=rev(levels(training$Class)), quiet=TRUE)
# Confusion Matrix
xgb.train.cm = confusionMatrix(as.factor(xgb.train.class), training$Class, positive = "1")
print(xgb.train.roc)
```
# Evaluate model performance in EARLI EHR derived
```{r Evaluate model performance in EARLI EHR, include=TRUE, echo=TRUE, message=TRUE}
# Predict and evaluate the model on EARLI EHR
xgb.pred = predict(xgb, earli_ehr, na.action = na.pass, type="prob")
xgb.pred = xgb.pred[,2] # Get whether or not we are predicting class 1
xgb.pred.class = ifelse(xgb.pred >= 0.50, 1, 0)
# Confusion Matrix
xgb.pred.cm = confusionMatrix(as.factor(xgb.pred.class), earli_ehr$Class, positive = "1")
# ROC
xgb.pred.roc = roc(predictor = xgb.pred, earli_ehr$Class, levels=rev(levels(earli_ehr$Class)), quiet=TRUE)
xgb.pred.roc.ci = ci(xgb.pred.roc)

earli.ehr.roc = glue("AUC: {round(as.numeric(xgb.pred.roc$auc), 2)}")
print(xgb.pred.roc)
print(xgb.pred.roc.ci)
```
# Evaluate model performance in EARLI Original, Hand-Curated
```{r Evaluate model performance in EARLI Curated, include=TRUE, echo=TRUE, message=TRUE}
# Predict and evaluate the model on EARLI Curated
xgb.pred = predict(xgb, earli_curated, na.action = na.pass, type="prob")
xgb.pred = xgb.pred[,2] # Get whether or not we are predicting class 1
xgb.pred.class = ifelse(xgb.pred >= 0.50, 1, 0)
# Confusion Matrix
xgb.pred.cm = confusionMatrix(as.factor(xgb.pred.class), earli_curated$Class, positive = "1")
# ROC
xgb.pred.roc = roc(predictor = xgb.pred, earli_curated$Class, levels=rev(levels(earli_curated$Class)), quiet=TRUE)
xgb.pred.roc.ci = ci(xgb.pred.roc)

earli.curated.roc = glue("AUC: {round(as.numeric(xgb.pred.roc$auc), 2)}")
print(xgb.pred.roc)
print(xgb.pred.roc.ci)
```